---
title: "Mapping noise across Accra, Ghana"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE}
suppressPackageStartupMessages({
library(stringr)
library(rgdal)
library(sp)  
library(dplyr)
library(tmap)
})
```

```{r , echo=FALSE}
day.long<-read.csv("/Users/sierraclark/Desktop/Sierras\ GitHub/Day.long.csv", header=TRUE)
cov.long<-read.csv(file="/Users/sierraclark/Desktop/Sierras\ GitHub/data_log_form_long.csv", header=TRUE)
long.cov<-merge(day.long, cov.long, by.x='ID', by.y='ID' ,all.x=TRUE)
```

```{r , echo=FALSE}
day.short<-read.csv("/Users/sierraclark/Desktop/Sierras\ GitHub/Day.short.csv", header=TRUE)
cov.short<-read.csv(file="/Users/sierraclark/Desktop/Sierras\ GitHub/data_log_form_short.csv", header=TRUE)
short.cov<-merge(day.short, cov.long, by.x='ID', by.y='ID' ,all.x=TRUE)
```

```{r , echo=FALSE, message=FALSE , warning=FALSE, results='hide', include=FALSE}
GAMA<-readOGR(dsn="/Users/sierraclark/Desktop/Sierras\ GitHub/spatial\ data/GAMA_boundV2.shp", layer="GAMA_boundV2")
Maj_roads<-readOGR(dsn="/Users/sierraclark/Desktop/Sierras\ GitHub/spatial\ data/Major_roads_OSM2019.shp", layer="Major_roads_OSM2019")
Maj_roads <- spTransform(Maj_roads, CRS("+init=epsg:32630")) # Re-project  to UTM 
Sec_roads<-readOGR(dsn="/Users/sierraclark/Desktop/Sierras\ GitHub/spatial\ data/Sec_ter_roads_OSM2019.shp", layer="Sec_ter_roads_OSM2019")
Sec_roads <- spTransform(Sec_roads, CRS("+init=epsg:32630")) # Re-project  to UTM 
GAMA_utm <- spTransform(GAMA, CRS("+init=epsg:32630")) # Re-project AMA to UTM 

Maj_roads <- Maj_roads[GAMA_utm, ] ## Clipping Road files to GAMA 
Sec_roads <- Sec_roads[GAMA_utm, ] ## Clipping Road files to GAMA 

water<-readOGR(dsn="/Users/sierraclark/Desktop/Sierras\ GitHub/spatial\ data/Lake_accra.shp", layer="Lake_accra")
water<- spTransform(water, CRS("+init=epsg:32630")) #re-project to UTM 
```

```{r , echo=FALSE}
coords <- cov.long[ , c("Lon", "Lat")]   # coordinates
data   <- cov.long[ , ]          # data
#crs    <- CRS("+init=epsg:32630") # projected 
crs    <- CRS("+init=epsg:4326")
# make the spatial points data frame object
long.sites <- SpatialPointsDataFrame(coords = coords,
                                     data = data, 
                                     proj4string = crs)

library(rgdal)
long_utm <- spTransform(long.sites, CRS("+init=epsg:32630")) # Re-project  to UTM 


coords <- cov.short[ , c("Lon", "Lat")]   # coordinates
data   <- cov.short[ , ]          # data
#crs    <- CRS("+init=epsg:32630") # projected 
crs    <- CRS("+init=epsg:4326")
# make the spatial points data frame object
short.sites <- SpatialPointsDataFrame(coords = coords,
                                     data = data, 
                                     proj4string = crs)

library(rgdal)
short_utm <- spTransform(short.sites, CRS("+init=epsg:32630")) # Re-project to UTM 
```

Knowing where the noisest areas are located in a city is important for policy, planning, and public awareness. 

In Accra, we took our minute-by-minute noise measurements, that were collected for either 7-days (136 rotating site locations) or a 1-year period (10 fixed site locations), and summarized the daily, day-time, and night-time metrics for each location. 

## LAeq24hr: Daily equivalent continuous sound levels (dBA) 
```{r , echo=FALSE, warning=FALSE, message=FALSE }
# Median LAeq24hr  
data.agg<-aggregate(short.cov$LAeq24hr, by=list(short.cov$ID), FUN=median)
data.agg<-plyr::rename(data.agg, c("Group.1"="ID", "x"="Median LAeq24hr (dBA)"))
short.spatial<- merge(short_utm, data.agg, by.x=c('ID'), by.y=c('ID'), all.x=TRUE) 

data.agg<-aggregate(long.cov$LAeq24hr, by=list(long.cov$ID), FUN=median)
data.agg<-plyr::rename(data.agg, c("Group.1"="ID", "x"="LAeq24hr"))
Fixed.data<- merge(long_utm, data.agg, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

short.spatial$num<- as.character(formatC(short.spatial$'Median LAeq24hr (dBA)',  big.mark = " ",  format = "f",   digits = 0))
Fixed.data$num<- as.character(formatC(Fixed.data$'LAeq24hr',  big.mark = " ",  format = "f",   digits = 0))
GAMA_utm$id<-as.character('GAMA')
Sec_roads$id<-as.character(' ')
Maj_roads$id<-as.character(' ')
water$id<-as.character(' ')

tmap_mode('view')
fig1 <-tm_shape(GAMA_utm) + tm_fill(col="grey90", alpha=0.4, id='id')  + tm_shape(Sec_roads) + tm_lines(col="grey77", legend.col.show = TRUE, id='id') +
  tm_shape(Maj_roads) + tm_lines(col="grey50", legend.col.show = TRUE, lwd=1, id='id') +
  tm_shape(short.spatial) + tm_bubbles( size=0.25, col="Median LAeq24hr (dBA)",  style="equal", palette="YlOrBr", id="num") +  
  tm_shape(Fixed.data) + tm_bubbles( size=0.7, col="LAeq24hr",shape=24, border.col="black", style="fixed", palette="YlOrBr",legend.col.show=FALSE, breaks=c(52, 57, 61, 66, 70 ,75), id="num") + 
  tm_shape(water) + tm_fill(col='cadetblue2', alpha=0.8, id='id') + 
  tm_layout(legend.position = c("right","bottom"), legend.format = list(scientific = TRUE, digits = 2), 
            legend.title.size = 1, legend.text.size=0.90, frame=FALSE)
fig1
```
*The larger circles indicate fixed sites (1 year)


## Lday: Day-time equivalent continuous sound levels (dBA) 
```{r , echo=FALSE, warning=FALSE, message=FALSE}
# Median Lday 
data.agg.day<-aggregate(short.cov$LDay, by=list(short.cov$ID), FUN=median)
data.agg.day$x<-as.numeric(data.agg.day$x)
data.agg.day<-plyr::rename(data.agg.day, c("Group.1"="ID", "x"="Median Lday (dBA)"))
short.spatial<- merge(short_utm, data.agg.day, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

data.agg.day<-aggregate(long.cov$LDay, by=list(long.cov$ID), FUN=median)
data.agg.day$x<-as.numeric(data.agg.day$x)
data.agg.day<-plyr::rename(data.agg.day, c("Group.1"="ID", "x"="Lday"))
Fixed.data<- merge(long_utm, data.agg.day, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

short.spatial$num<- as.character(formatC(short.spatial$'Median Lday (dBA)',  big.mark = " ",  format = "f",   digits = 0))
Fixed.data$num<- as.character(formatC(Fixed.data$'Lday',  big.mark = " ",  format = "f",   digits = 0))
GAMA_utm$id<-as.character('GAMA')
Sec_roads$id<-as.character(' ')
Maj_roads$id<-as.character(' ')
water$id<-as.character(' ')

fig2 <-tm_shape(GAMA_utm) + tm_fill(col="grey90", alpha=0.4, id='id') + tm_shape(Sec_roads) + tm_lines(col="grey77", legend.col.show = TRUE, id='id') +
  tm_shape(Maj_roads) + tm_lines(col="grey50", legend.col.show = TRUE, lwd=1, id='id') + 
  tm_shape(short.spatial) + tm_bubbles( size=0.25, col="Median Lday (dBA)", style="equal", palette="Reds", 
                                        id="num") +  
  tm_shape(Fixed.data) + tm_bubbles( size=0.7, col="Lday",shape=24, border.col="grey30", style="fixed", palette="Reds",legend.col.show=FALSE , breaks=c(53, 58, 62, 67, 71, 76), id="num") + 
  tm_shape(water) + tm_fill(col='cadetblue2', alpha=0.8) + 
  tm_layout(legend.position = c("right","top"), legend.format = list(scientific = TRUE, digits = 2), 
            legend.title.size = 1, legend.text.size=0.90, frame=FALSE) 
fig2
```
*The larger circles indicate fixed sites (1 year)

## Lnight: Night-time equivalent continuous sound levels (dBA) 
```{r , echo=FALSE, warning=FALSE, message=FALSE}
# Median LNight
data.agg.night<-aggregate(short.cov$LNight, by=list(short.cov$ID), FUN=median)
data.agg.night$x<-as.numeric(data.agg.night$x)
data.agg.night<-plyr::rename(data.agg.night, c("Group.1"="ID", "x"="Median Lnight (dBA)"))
short.spatial<- merge(short_utm, data.agg.night, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

data.agg.night<-aggregate(long.cov$LNight, by=list(long.cov$ID), FUN=median)
data.agg.night$x<-as.numeric(data.agg.night$x)
data.agg.night<-plyr::rename(data.agg.night, c("Group.1"="ID", "x"="Lnight"))
Fixed.data<- merge(long_utm, data.agg.night, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

short.spatial$num<- as.character(formatC(short.spatial$'Median Lnight (dBA)',  big.mark = " ",  format = "f",   digits = 0))
Fixed.data$num<- as.character(formatC(Fixed.data$'Lnight',  big.mark = " ",  format = "f",   digits = 0))
GAMA_utm$id<-as.character('GAMA')
Sec_roads$id<-as.character(' ')
Maj_roads$id<-as.character(' ')
water$id<-as.character(' ')

fig3 <-tm_shape(GAMA_utm) + tm_fill(col="grey90", alpha=0.4, id='id') +  tm_shape(Sec_roads) + tm_lines(col="grey77", legend.col.show = TRUE, id='id') +
  tm_shape(Maj_roads) + tm_lines(col="grey50", legend.col.show = TRUE, lwd=1, id='id') + 
  tm_shape(short.spatial) + tm_bubbles( size=0.25, col="Median Lnight (dBA)",  style="equal", palette="BuPu",
                                       id="num") +  
  tm_shape(Fixed.data) + tm_bubbles( size=0.7, col="Lnight",shape=24, border.col="grey30",  style="fixed", palette="BuPu",legend.col.show=FALSE, breaks=c(46,52, 57, 62, 67, 72 ), id="num" ) + 
  tm_shape(water) + tm_fill(col='cadetblue2', alpha=0.8) + 
  tm_layout(legend.position = c("right","bottom"), legend.format = list(scientific = TRUE, digits = 2), legend.title.size = 1, legend.text.size=0.90, frame=FALSE) + 
  tm_add_legend(type = c("line"), labels = c("Major roads", "Secondary roads"), 
                col = c("grey30", "grey60"), lty = 1, lwd=c(1.5, 1), title="") + 
  tm_scale_bar(width=0.15, size=1.2)
fig3
```
* The larger circles indicate fixed sites (1 year)


## IR24hr: Daily Intermittency Ratios (%) 
(What percentage of the total sound energy is from intermittent events (e.g., Ambulance siren or an airplane flying overhead)) 
```{r ,echo=FALSE , warning=FALSE, message=FALSE}
# Median IR 
data.agg.ir<-aggregate(short.cov$IR, by=list(short.cov$ID), FUN=median)
data.agg.ir$x<-as.numeric(data.agg.ir$x)
data.agg.ir<-plyr::rename(data.agg.ir, c("Group.1"="ID", "x"="Median IR24hr (%)"))
short.spatial<- merge(short_utm, data.agg.ir, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

data.agg.ir<-aggregate(long.cov$IR, by=list(long.cov$ID), FUN=median)
data.agg.ir$x<-as.numeric(data.agg.ir$x)
data.agg.ir<-plyr::rename(data.agg.ir, c("Group.1"="ID", "x"="IR"))
Fixed.data<- merge(long_utm, data.agg.ir, by.x=c('ID'), by.y=c('ID'),  all.x=TRUE) 

short.spatial$num<- as.character(formatC(short.spatial$'Median IR24hr (%)',  big.mark = " ",  format = "f",   digits = 0))
Fixed.data$num<- as.character(formatC(Fixed.data$'IR',  big.mark = " ",  format = "f",   digits = 0))
GAMA_utm$id<-as.character('GAMA')
Sec_roads$id<-as.character(' ')
Maj_roads$id<-as.character(' ')
water$id<-as.character(' ')

fig4 <-tm_shape(GAMA_utm) + tm_fill(col="grey90", alpha=0.4, id='id') +  tm_shape(Sec_roads) + tm_lines(col="grey77", legend.col.show = TRUE, id='id') +
  tm_shape(Maj_roads) + tm_lines(col="grey50", legend.col.show = TRUE, lwd=1, id='id') + 
  tm_shape(short.spatial) + tm_bubbles( size=0.25, col="Median IR24hr (%)",  style="equal", palette="GnBu", 
                                        id="num") +  
  tm_shape(Fixed.data) + tm_bubbles( size=0.7, col="IR",shape=24, border.col="grey30", style="fixed", palette="GnBu",legend.col.show=FALSE ,breaks=c(15, 30, 44, 58, 72, 86), id="num") + 
  tm_shape(water) + tm_fill(col='cadetblue2', alpha=0.8) + 
  tm_layout(legend.position = c("right","bottom"), 
            legend.format = list(scientific = TRUE, digits = 2), legend.title.size = 1, legend.text.size=0.90, frame=FALSE) 
fig4
```
* The larger circles indicate fixed sites (1 year)

Figures are adapted from the open access peer-reviewed published paper Clark et al. Scientific Reports. 2021.




